---
title: "Multilinear Regression"
author: "Matthew Bardoe & Carey Kopeikin"
date: "4/15/2020"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```



## Your Turn

Load up the dataset for Prices for Used Camrys. Follow the steps to create a multilinear model to estimate the price of a Used Camry. 

```{R}
camry <- read.csv("Camry.csv")
head(camry)
```


**Question** Create a table show the correlations of all the numeric variables.

```{r}
cor(camry[c(-2, -3, -4)])
```




**Question** Create a plot showing scatterplots of all the numeric variables.

```{r}
plot(camry[c(-2, -3, -4)])
```



**Question** Create a simple linear model predicting price from the mileage. Find the percent of the variation of the price that is explained by the variation in the mileage.


```{r}
plot( camry$Price ~ camry$Mileage, 
     main = "Scatter Plot of Price vs. Mileage",
     xlab = "Mileage",
     ylab = "Price",
     col = "red",  
     pch = 18,
     cex = 1.5)

```


**Question** Write down the linear model.
```{r}
linearMod.price.mileage <- lm(Price ~ Mileage, data = camry)
print(linearMod.price.mileage)

summary(linearMod.price.mileage)
```

*Predicted price = -.075(miles) + 15,830



**Question** Explain what the model written above means in context.
*The three asterisks means that the relationship between price and mileage is not just due to random variation.*

*For each additional mile I would expect the price to go down by $.07. If a car had 0 miles I would predict that it would cost 15,000.*

**Question** Create the best multi-linear model you can predicting price from all other numeric variables that are significant to the price. 

```{r}
multiMod.price.mileage.year <- lm(Price ~ Mileage + ModelYear + MPG_Highway, data=camry)

summary(multiMod.price.mileage.year)

```


```{r}
multiMod.price.mileage.year <- lm(Price ~ Mileage + ModelYear, data=camry)

summary(multiMod.price.mileage.year)

```


**Question** Write down the multi linear model.
*Predicted Price = -.02889(Mileage) + 615.9(ModelYear) -1,227,000



**Question** Explain what the model written above means in context.
*This means that for each additional model year, we expect the price will go up by $615.9 holding all other variables constant and for each additional mile on the car we expect the price  will decrease by 0.02889 holding all other variables constant.*



**Question** Find the percent of the variation of the price that is explained by this model.
*It explains 78.5% of car prices.*



**Question** How much better is the multilinear model than the simple linear model?
*The simple model explained 65.44% of prices, but the multilinear model explains 78.5%, so the second one is .785/.6544 = 1.2 times better.*




**Question** Predict the price of 3 cars: A 2013 Camry with 32,000 miles, a 2010 Camry with 89000, and a 2016 Camry with 90000 miles.


```{r}
predict.data.frame <- data.frame( Mileage = c(32000, 89000, 90000 ), ModelYear = c(2013, 2010, 2016) )

predict.data.frame$Predicted.price <- predict(multiMod.price.mileage.year, predict.data.frame)

predict.data.frame


```

```{r}
-.02889*(32000) + 615.9*(2013) -1227000

```


```{r}
-.02889*(Mileage) + 615.9*(ModelYear) -1227000

```


















