---
title: "Regression Lab Practice"
author: "Matthew Bardoe & Carey Kopeikin"
date: "2/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

## Regression Lab Practice

In the up coming lab you will be asked to choose from several datasets. You will be asked to determine some variable within that dataset to try and predict. Then in a document show your process to finding that model. Finally you will be given a few rows of data and you are to compare your predictions in those rows with the true answers in the data.

This assignment is to practice that process with the following dataset.

```{r}

kc.house=read.csv("../../data/kc_house_data.csv")
head(kc.house)
str(kc.house)

```

<https://www.kaggle.com/swathiachath/kc-housesales-data>

## Instructions

-   Explore the data, and decide on a variable to predict using linear or multilinear regression

-   Use the techniques that you have learned in class to create an equation for that variable, and make sure that your equation explains as much as you can, but does not overfit the data. Show your work and comment on the choices that you make. Tell the story of how you came to this model.

-   Use your model to predict the value of the chosen variable for this group that are not part of the original dataset.

Instructions that I will give you for this practice.

-   Be sure not to include variables that have no stars
-   Try to include both quantitative and categorical data
-   Try to make the R-squared as high as possible
-   Make sure that you make a residual plot to test your model.

### Exploration

Discuss your method for finding the model. Explain what you are looking for and how you are going about creating your model. Explain why you are choosing the variables and how you can tell how good your model is.

### Model Creation

Write down the equation of your model and explain what each of the numbers means.

### Prediction

Use your model to predict the values of the houses in the `kc_predict.csv` data file. And then comment on how well your model did.

```{r}
kc_predic=read.csv("../../data/kc_predict.csv")
```

