---
title: "Linear Regression Practice"
author: "Carey Kopeikin"
date: "2/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```


Load up the data set Indian.Ag

```{r}
iag.data <- read.csv("IndianAg.csv")
```

Create a scatter plot of Cost of Cultivation vs. Yield. Make sure to put the explanatory and response variables on the correct axis.

```{r}
plot(iag.data$Yield ~ iag.data$Cost.of.Cultivation)

```

Is it appropriate to use a linear model to describe the association between the variables? Why or why not?

*Your answer here*
Yes both variables are quantitative and the data appears linear


Create a linear model for the two variables.

```{r}
lin.mod.Yield.Coc <- lm( Yield ~ Cost.of.Cultivation, data = iag.data)
lin.mod.Yield.Coc
```


Find the predicted values and the residuals and add them to the data frame. 

```{r}
iag.data$Predictions <- predict(lin.mod.Yield.Coc)
iag.data$Residuals <- resid(lin.mod.Yield.Coc)
```


Make a graph of the predicted values and the residuals.

```{r}
plot(iag.data$Residuals ~ iag.data$Predictions)

abline(0,0)

```


Do you still think a linear model is appropriate? Why?

Yes the residual plot is clould like which confirms that the original data was linear.

Find the summary of the linear model.

```{r}
summary(lin.mod.Yield.Coc)
```

Write the equation of the model

predicted Yield = .0010814 * Cost.of.Cultivation + 1.2908493

Could the association between Cost of Cultivation and Yield be due to random variation? Why or why not?

No as there are 3 stars next to the cost of cultivation. This means that the likelihood that the relationship we found between the the two variables due to coincidence is very small. 


How good is the model at explaining the variation in Yield?

Not great as cost of cultivation only explains 26.12% of the variation in Yeild.

Part 2

Create a scatter plot of Extra Costs vs. Yield. Make sure to put the explanatory and response variables on the correct axis.

```{r}


```

Is it appropriate to use a linear model to describe the association between the variables? Why or why not?

*Your answer here*



Create a linear model for the two variables.

```{r}
```


Find the predicted values and the residuals and add them to the data frame. 

```{r}

```


Make a graph of the predicted values and the residuals.

```{r}


```


Do you still think a linear model is appropriate? Why?



Find the summary of the linear model.

```{r}

```

Write the equation of the model



Could the association between Extra Costs and Yield be due to random variation? Why or why not?





How good is the model at explaining the variation in Yield?


