---
title: "Day 10 Linear Regression Part 3"
author: "Carey Kopeikin"
date: "1/9/2022 rev 1/5/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Linear Regression Part 3

What you will learn in this lesson:

* Interpreting residual plots


#Example 1
Upload the data set ```nonlinear.examples.csv``` and call it ```examples.data```

```{r}
examples.data <- read.csv("../../data/nonlinear.examples.csv")
head(examples.data)
```


Create a scatterplot with labels with Var1 as the explanatory variable and Var2 as the response variable. 

```{r}

```

Describe the shape, direction, and strength of the association:
*Your answer here.*



Is it appropriate to use correlation to talk about the relationship between these variables? Explain why or why not.

*Your answer here.*



Find the correlation
```{r}

```


Create a linear model and print out the results.

```{r}

```

Make a scatterplot that includes the line of best fit.

```{r}

```

Find the predicted values and the residuals and add them to the data frame. 

```{r}

```


Make a graph of the predicted values and the residuals.

```{r}


```


Do you still think a linear model is appropriate? Why?






## Example 2

Upload the data set ```nonlinear.examples.csv``` and call it ```examples.data```

```{r}
examples.data <- read.csv("../../data/nonlinear.examples.csv")
```


Create a scatterplot with labels with Var1 as the explanatory variable and Var3 as the response variable. 

```{r}

```

Describe the shape, direction, and strength of the association:
*Your answer here.*



Is it appropriate to use correlation to talk about the relationship between these variables? Explain why or why not.

*Your answer here.*



Find the correlation
```{r}

```


Create a linear model and print out the results.

```{r}

```

Make a scatterplot that includes the line of best fit.

```{r}

```

Find the predicted values and the residuals and add them to the data frame. 

```{r}

```


Make a graph of the predicted values and the residuals.

```{r}


```


Do you still think a linear model is appropriate? Why?



```{r}

```


