---
title: "Multilinear Regression"
author: "Matthew Bardoe & Carey Kopeikin"
date: "4/15/2020"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```



## Your Turn

Load up the dataset for Prices for Used Camrys. Follow the steps to create a multilinear model to estimate the price of a Used Camry. 

```{R}
camry <- read.csv("Camry.csv")
head(camry)
```


**Question** Create a table that shows the correlations of all the numeric variables.

```{r}

```




**Question** Create a plot showing scatterplots of all the numeric variables.

```{r}

```



**Question** Create a simple linear model predicting price from the mileage. Find the percent of the variation of the price that is explained by the variation in the mileage.


```{r}


```


**Question** Write down the linear model.




**Question** Explain what the model written above means in context.




**Question** Create the best multi-linear model you can predicting price from all other numeric variables that are significant to the price. 

```{r}


```


**Question** Write down the multi linear model.




**Question** Explain what the model written above means in context.




**Question** Find the percent of the variation of the price that is explained by this model.




**Question** How much better is the multilinear model than the simple linear model?





**Question** Predict the price of 3 cars: A 2013 Camry with 32,000 miles and 25mpg highway, a 2010 Camry with 89000 and 21mpg highway, and a 2016 Camry with 90000 miles and 26mpg highway.


```{r}



```






















